Parallel training by Pipeline Execution In AIRCoP

### 2021-12-19 report/paper.tex
complete script to generate plot from csv raw data
Second round Spelling/Grammar Pass

Operations are connected through let-bindings. Furthermore, we use annotation of the formal syntax to represent different parallelism strategies. We develop annotation systems for both Tensor, Operation, and Module. Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition, and Module annotation represents the number of micro-batches. Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.

@@

Stacked Pipeline Parallelism will introduce little communication overhead because it will only transfer data at the module boundaries. This feature of Stacked Pipeline Parallelism makes it suitable both for systems with high-speed interconnections and weakly connected systems. \par

@@

$G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass. All the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase because our source language does not allow mutation. We only need to write the versioned lift node at the end of the forward pass, and the Generation phase can generate the correct training loop.

@@

The typed frontend source language of AIRCoP only contains the sequential forward computation of the model and annotations. The source language does not mutate tensors. We use overloading of the typed frontend objects and reflection to generate the corresponding IR of the source program. We put the explicit TensorType and Annotations into the RHS of the IR to make it persistent and used in later transformation. \par

@@

We statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the loop to optimize memory management. We currently use code-motion techniques to lift the cudaMalloc out of the training loop and pre-allocate the memory to optimize memory management. This approach could significantly improve performance because cudaMalloc/Free often occupies excessive time.\par

@@

  \includegraphics[scale=0.5]{pipelineparallelruntime.png}

@@

  \includegraphics[scale=0.5]{pipelineparallelspeedup.png}

@@

However, we find that 2 modules on Stacked Pipeline Parallelism with each module running at Data Parallelism on 2 GPUs, i.e., Mixed Parallelism, still perform a little worse than Data Parallelism on 4 GPUs. This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible. However, the cross-module dependencies for micro-batch will still cause some bubble time. As we mentioned before, the ideal network topologies for Pipeline Parallelism are an extensive training system consisting of multiple GPU clusters connected through the network. Then we can use Data Parallelism inside each cluster and Pipeline Parallelism among clusters. We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could deduce that Pipeline Parallelism mixed with Data Parallelism will perform much better than only using Data Parallelism under this scenario because Pipeline Parallelism could significantly reduce the communication overhead by restricting the inter-cluster communication to only happen at module boundaries.\par

>>>>

Operations are connected through let-bindings. We use annotation of the formal syntax to represent different parallelism strategies. We develop annotation systems for both Tensor, Operation, and Module. Tensor and Operation annotation represents Spatial Partition by indicating the dimension to partition, and Module annotation represents the number of micro-batches. Users only need to write the sequential forward computation of the Machine Learning model, and the correct final training loop will be automatically computed through a series of transformations.

@@

Stacked Pipeline Parallelism will introduce little communication overhead because it will only transfer data at the module boundaries. This feature of Stacked Pipeline Parallelism makes it suitable both for systems with high-speed interconnections as well as weakly connected systems. \par

@@

$G_w$ generates the weight and their gradient and momentum, $G_l$ generate an array of lift nodes for different micro-batches, $G_f$ generates forward pass and write the correct value to the versioned lift variables, $G_b$ generates the backward pass and read from the versioned lift variable, $G_o$ generates the optimization pass. All the mutation operations (Assign/Accumulate/Optimize) are automatically generated in the Generation Phase because our source language does not allow mutation, so we only need to write the versioned lift nodes at the end of the forward pass.

@@

The typed frontend source language of AIRCoP only contains the sequential forward computation of the model and annotations. The source language does not mutate tensors. We use overloading of the typed frontend objects and reflection to generate the corresponding IR of the source program. We put the explicit TensorType and Annotations into the RHS of the IR to make them persistent and use them in later transformation passes. \par

@@

We statically allocate GPU memory before the training loop and avoid time-exhausting cudaMalloc/cudaFree inside the loop to optimize memory management. We currently use code-motion techniques to lift the cudaMalloc out of the training loop and pre-allocate the memory to optimize memory management. This approach could significantly improve performance because cudaMalloc/Free often occupies excessive amount of time.\par

@@

  \includegraphics[scale=0.5]{pipelineparallelruntime.pdf}

@@

  \includegraphics[scale=0.5]{pipelineparallelspeedup.pdf}

@@

However, we find that 2 modules on Stacked Pipeline Parallelism with each module running at Data Parallelism on 2 GPUs, i.e., Mixed Parallelism, still perform a little worse than Data Parallelism on 4 GPUs. This result is anticipated because we perform our evaluation on a single GPU cluster with very high-speed inter-GPU connections, which makes the frequent tensor communication (ALLReduce) overhead introduced by Data Parallelism negligible. However, the cross-module dependencies of micro-batches in Pipeline Parallelism will still cause some bubble time. As we mentioned before, the ideal network topologies for Pipeline Parallelism are an extensive training system consisting of multiple GPU clusters connected through the network. Then we can use Data Parallelism inside each cluster and Pipeline Parallelism among clusters. We currently could not perform evaluation under this ideal scenario due to limited time and hardware resources, but we could deduce that Pipeline Parallelism mixed with Data Parallelism will perform much better than only using Data Parallelism under this scenario because Pipeline Parallelism could significantly reduce the communication overhead by restricting the inter-cluster communication to only happen at module boundaries.\par
